Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2021-06-17 thirdPaper/draft/typeConeAssociahedra.tex
massage

Our assumptions allow for a well-defined notion of index:~$\ind_T(X)=[T_0^X]-[T_1^X]\in\kzero{\add T}$.

@@

Fix an essential complete simplicial fan~$\Fan$ in~$\R^n$, with an (arbitrary) indexing of its rays by $[N]\eqdef \{1,\dots,N\}$. Let~$\b{G}$ be the $N \times n$-matrix whose rows are representative vectors of the rays of~$\Fan$. Let~$\b{K}$ be a $(N-n) \times N$-matrix that spans the left kernel of~$\b{G}$ (\ie $\b{K}\b{G} = 0$ and $\rank(\b{K})=N-n$). For any height vector~$\b{h} \in \R^N$, we define the polytope
\[
P_\b{h} \eqdef \set{\b{x} \in \R^n}{\b{G}\b{x} \le \b{h}},
\]
where $\b{G}\b{x} \le \b{h}$ is the standard shorthand for the corresponding system of inequalities.

@@

\begin{example}
\label{exm:typeCone}
\enlargethispage{.5cm}
Consider the $2$-dimensional fan~$\Fan$ depicted in \cref{fig:typeCone} (left) (we will encounter this fan again in \cref{exm:typeConeCoarsenings,exm:typeConeMinkowskiSum,exm:typeConeAsso}).
It has five rays labeled~$1, \dots, 5$ and five maximal cones labeled~$a,\dots, e$.
The matrices $\b{G}$ and $\b{K}$ are (for a possible choice of $\b{K}$): 
\begin{align*}
\transpose{\b{G}}&=\begin{bmatrix} 1 & 0 & -1& -1 & 0 \\ 0 & 1 & 1 & 0 &-1 \end{bmatrix},&
\b{K}&=\begin{bmatrix} 1 & 0 & 0 & 1 & 0\\ 0 & 1 & 0 & 0 & 1 \\ 1& 0& 1& 0 & 1  \end{bmatrix}.
\end{align*}
Its type cone~$\typeCone(\Fan)$ lies in~$\R^5$, but has a $2$-dimensional lineality space.
The five pairs of adjacent maximal cones of~$\Fan$ give rise to following five defining inequalities for~$\typeCone(\Fan)$:
\begin{gather*}
H^>_{ab}: \; \b{h}_1 + \b{h}_3 - \b{h}_2 > 0
\qquad
H^>_{bc}: \; \b{h}_2 + \b{h}_4 - \b{h}_3 > 0
\qquad
H^>_{cd}: \; \b{h}_3 + \b{h}_5 - \b{h}_4 > 0
\\
H^>_{de}: \; \b{h}_1 + \b{h}_4 > 0
\qquad
H^>_{ae}: \; \b{h}_2 + \b{h}_5 > 0,
\end{gather*}
where $H^>_{ij}$ denotes the halfspace defined by the inequality corresponding to the two adjacent maximal cones~$i$ and~$j$.
Note that the inequalities~$H^>_{ae}$ and~$H^>_{de}$ are redundant.
For example, the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$ belongs to the type cone~$\typeCone(\Fan)$, and the corresponding polytope is represented in \cref{fig:typeCone} (middle).
To represent this type cone, it suffices to represent~$\b{K}\typeCone(\Fan)$ which is an essential simplicial cone in~$\R^3$, given as the intersection of the open halfspaces:
\begin{gather*}
\b{K}H^>_{ab}: \; {x} - {y} +2 {z} > 0
\qquad
\b{K}H^>_{bc}: \; {x} + {y} - {z} > 0
\qquad
\b{K}H^>_{cd}: \; -{x} + {y} +2 {z} > 0
\\
\b{K}H^>_{de}: \; 2{x} + {z} > 0
\qquad
\b{K}H^>_{ae}: \; 2{y} + {z} > 0.
\end{gather*}
One can further reduce the dimension by intersecting with a transversal hyperplane to get a $2$-dimensional representation.
This is the red triangle depicted in \cref{fig:typeCone} (right).

\begin{figure}[htpb]
	\capstart
	\centerline{{\includegraphics[scale=1]{typeConeFan}} \quad {\includegraphics[scale=1]{typeConePolytope}} \quad \raisebox{.5cm}{\includegraphics[scale=.45]{typeCone}}}
	\caption{A $2$-dimensional fan~$\Fan$ with five rays $1, \dots, 5$ and five maximal cones~$a, \dots, e$ (left), its polytopal realization corresponding to the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, 1, 1, \nicefrac{5}{4})$ (middle), and the intersection of $\b{K}\typeCone(\Fan)$ with a hyperplane to get a $2$-dimensional slice (right).}
	\label{fig:typeCone}
\end{figure}
\end{example}

@@

\enlargethispage{.5cm}
We now discuss the connection between convex combinations in the type cone and Minkowski sums of polytopes.
Recall that the \defn{Minkowski sum} of two polytopes~$P$ and~$Q$ is the polytope $P + Q \eqdef \set{p + q}{p \in P\text{ and } q \in Q}$.

@@

\begin{example}
\label{exm:typeConeMinkowskiSum}
\enlargethispage{.5cm}
Consider the $2$-dimensional fan~$\Fan$ of \cref{fig:typeCone}\,(left) whose type cone has been described in \cref{exm:typeCone} and represented in \cref{fig:typeCone}\,(right).
As illustrated in \cref{fig:typeConeCoarsenings}, the rays of the closed type cone~$\ctypeCone(\Fan)$ are directed by the height vectors~$(1,2,1,2,1)$ (corresponding to a triangle), $(0,1,1,0,1)$ (corresponding to a vertical segment) and~$(1,0,1,1,0)$ (corresponding to a horizontal segment).
Therefore, any polytope in~$\ctypeCone(\Fan)$ is (up to translation) a positive Minkowski combination of a triangle and two edges.
\cref{fig:typeConeMinkowskiSum} illustrates this property for the polytope of \cref{fig:typeCone}\,(middle) corresponding to the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$.
Note that~$\Fan$ is the normal fan of J.-L.~Loday's $2$-dimensional associahedron, whose realization as a Minkowski sum of faces of a simplex was given by A.~Postnikov in~\cite{Postnikov}.

\begin{figure}[htpb]
	\capstart
	\centerline{
	\begin{tabular}{c@{\;$=$\;}c@{\,$+$\,}c@{\,$+$\,}c@{\,}l@{\,}l}
	$(\frac{1}{2}, \frac{3}{4}, \frac{5}{4}, 1, \frac{5}{4})$ & $\frac{1}{6}(1,2,1,2,1)$ & $\frac{3}{4}(0,1,1,0,1)$ & $\frac{1}{2}(1,0,1,1,0)$ & $-\,\frac{1}{6}(1,0,-1,-1,0)$ & $-\,\frac{1}{3}(0,1,1,0,-1)$ \\
	\raisebox{-0.5\height}{\includegraphics[scale=.5]{typeConePolytope}} & \raisebox{-0.5\height}{\includegraphics[scale=.5]{typeConeMinkowskiSum1}} & \raisebox{-0.5\height}{\includegraphics[scale=.5]{typeConeMinkowskiSum2}} & \raisebox{-0.5\height}{\includegraphics[scale=.5]{typeConeMinkowskiSum3}} & $+$ translation.
	\end{tabular}
	}
	\caption{The polytope of \cref{fig:typeCone}\,(middle) is a Minkowski sum of a triangle and two edges.}
	\label{fig:typeConeMinkowskiSum}
\end{figure}
\end{example}

@@

\enlargethispage{.1cm}
We now focus on the normal fan of~$\Asso[n]$.
Since~$\Asso[n]$ lives in a hyperplane of~$\R^{n+1}$, its normal fan has a one-dimensional lineality.
Let~$\HH \eqdef \bigset{\b{x} \in \R^{n+1}}{\sum_{\ell \in [n+1]} \b{x}_\ell = 0}$ and~$\pi : \R^{n+1} \to \HH$ denote the orthogonal projection. We denote by~$\gvectorFan[n]$ the intersection of the normal fan of~$\Asso[n]$ with~$\HH$, which is an essential complete simplicial fan.
The following facts are illustrated in \cref{fig:lodayFans} in dimension~$2$ and~$3$.

@@

% In view of \cref{def:gvectorCA}, it is convenient to express the $\b{g}$-vectors in the basis given by $\gvector{0,i+1}$ for $i\in [n]$. This is what we do in \cref{fig:lodayFans} and \cref{exm:typeConeAsso}.
Note that in \cref{fig:lodayFans} and \cref{exm:typeConeAsso} we express the $\b{g}$-vectors in the basis given by $\gvector{0,i+1}$ for $i\in [n]$ (to be coherent with the upcoming \cref{def:gvectorCA}).

\begin{figure}[h]
	\capstart
	\centerline{\raisebox{.5cm}{\includegraphics[scale=.45]{lodayFan2}}\hspace{1.5cm}\includegraphics[scale=.45]{lodayFan3}}
	\caption{The normal fans~$\gvectorFan[2]$ and~$\gvectorFan[3]$ of J.-L.~Loday's associahedra. As the rightmost fan is $3$-dimensional, we intersect it with the sphere and stereographically project it from the direction~$(-1,-1,-1)$.}
	\label{fig:lodayFans}
\end{figure}

@@

From the linear dependencies of \cref{prop:exchangeablePairsAsso}, we obtain a redundant description of the type cone of the fan~$\gvectorFan[n]$.
To simplify the presentation, we embed the type cone in a larger space, adding dummy variables for the boundary edges of the polygon.

@@

@@

@@

\[
\begin{array}{r|cccccccccc}
\text{diagonals} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB1}}  & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB2}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB3}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB4}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB5}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB6}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB7}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB8}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB9}} \\[.3cm]
\text{$\b{g}$-vectors} & \compactVectorT{1}{0}{0} & \compactVectorT{0}{1}{0} & \compactVectorT{0}{0}{1} & \compactVectorT{-1}{1}{0} & \compactVectorT{0}{-1}{1} & \compactVectorT{-1}{0}{1} & \compactVectorT{-1}{0}{0} & \compactVectorT{0}{-1}{0} & \compactVectorT{0}{0}{-1} \\[.6cm]
\text{facet}		& 1 & -1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & \red \circled{A} \\
\text{defining} 	& 0 & 1 & -1 & -1 & 1 & 0 & 0 & 0 & 0 & \red \circled{B} \\
\text{inequalities}	& 0 & 0 & 1 & 0 & -1 & 0 & 1 & 0 & 0 & \red \circled{C} \\
					& 0 & 0 & 0 & 1 & -1 & 1 & 0 & 0 & 0 & \red \circled{D} \\
					& 0 & 0 & 0 & 0 & 1 & -1 & -1 & 1 & 0 & \red \circled{E} \\
					& 0 & 0 & 0 & 0 & 0 & 1 & 0 & -1 & 1 & \red \circled{F} \\[.2cm]
\end{array}
\]
\medskip

@@

\begin{figure}[h]
	\capstart
    \begin{adjustbox}{center}
        \begin{tikzpicture}
        	\matrix (m) [matrix of math nodes, row sep=.35cm, column sep=.2cm, nodes={anchor=center, align=center, inner sep=0pt}]{
        		& \;\includegraphics[scale=.7]{diagonalA2} & \node (b) {\red \circled{B}}; & \;\includegraphics[scale=.7]{diagonalA4} & \\
        		\;\includegraphics[scale=.7]{diagonalA1} & \node (a) {\red \circled{A}}; & \;\includegraphics[scale=.7]{diagonalA3} & \node (c) {\red \circled{C}}; & \;\includegraphics[scale=.7]{diagonalA5} \\};
        	\draw[->] (m-2-1) -- (m-1-2);
        	\draw[->] (m-1-2) -- (m-2-3);
        	\draw[->] (m-2-3) -- (m-1-4);
        	\draw[->] (m-1-4) -- (m-2-5);
        	\draw[red] (a) -- (m-2-1);
        	\draw[red] (a) -- (m-2-3);
        	\draw[red] (a) -- (m-1-2);
        	\draw[red] (b) -- (m-1-2);
        	\draw[red] (b) -- (m-1-4);
        	\draw[red] (b) -- (m-2-3);
        	\draw[red] (c) -- (m-2-3);
        	\draw[red] (c) -- (m-2-5);
        	\draw[red] (c) -- (m-1-4);
        \end{tikzpicture}
        \quad
        \begin{tikzpicture}
        	\matrix (m) [matrix of math nodes, row sep=.35cm, column sep=.2cm, nodes={anchor=center, align=center, inner sep=0pt}]{
        		&& \;\includegraphics[scale=.7]{diagonalB3} & \node (c) {\red \circled{C}}; & \;\includegraphics[scale=.7]{diagonalB7} && \\
        		& \;\includegraphics[scale=.7]{diagonalB2} & \node (b) {\red \circled{B}}; & \;\includegraphics[scale=.7]{diagonalB5} & \node (e) {\red \circled{E}}; & \;\includegraphics[scale=.7]{diagonalB8} & \\
        		\;\includegraphics[scale=.7]{diagonalB1} & \node (a) {\red \circled{A}}; & \;\includegraphics[scale=.7]{diagonalB4} & \node (d) {\red \circled{D}}; & \;\includegraphics[scale=.7]{diagonalB6} & \node (f) {\red \circled{F}}; & \;\includegraphics[scale=.7]{diagonalB9} \\};
        	\draw[->] (m-3-1) -- (m-2-2);
        	\draw[->] (m-2-2) -- (m-1-3);
        	\draw[->] (m-2-2) -- (m-3-3);
        	\draw[->] (m-1-3) -- (m-2-4);
        	\draw[->] (m-3-3) -- (m-2-4);
        	\draw[->] (m-2-4) -- (m-1-5);
        	\draw[->] (m-2-4) -- (m-3-5);
        	\draw[->] (m-1-5) -- (m-2-6);
        	\draw[->] (m-3-5) -- (m-2-6);
        	\draw[->] (m-2-6) -- (m-3-7);
        	\draw[red] (a) -- (m-3-1);
        	\draw[red] (a) -- (m-3-3);
        	\draw[red] (a) -- (m-2-2);
        	\draw[red] (b) -- (m-2-2);
        	\draw[red] (b) -- (m-2-4);
        	\draw[red] (b) -- (m-1-3);
        	\draw[red] (b) -- (m-3-3);
        	\draw[red] (c) -- (m-1-3);
        	\draw[red] (c) -- (m-1-5);
        	\draw[red] (c) -- (m-2-4);
        	\draw[red] (d) -- (m-3-3);
        	\draw[red] (d) -- (m-3-5);
        	\draw[red] (d) -- (m-2-4);
        	\draw[red] (e) -- (m-2-4);
        	\draw[red] (e) -- (m-2-6);
        	\draw[red] (e) -- (m-1-5);
        	\draw[red] (e) -- (m-3-5);
        	\draw[red] (f) -- (m-3-5);
        	\draw[red] (f) -- (m-3-7);
        	\draw[red] (f) -- (m-2-6);
        \end{tikzpicture}
    \end{adjustbox}
	\caption{The facet-defining inequalities of the type cone~$\typeCone \big( \gvectorFan[n] \big)$ correspond to the flips described in \cref{prop:extremalExchangeablePairsAsso}. See \cref{subsec:typeConeCA,sec:clusterCategories} for a representation theoretic \mbox{interpretation}.}
	\label{fig:labelFacetDefiningInequalititiesAsso}
\end{figure}
\end{example}

@@

\enlargethispage{.2cm}
Combining \cref{coro:simplicialTypeCone,coro:simplicialTypeConeAsso} and \cref{prop:extremalExchangeablePairsAsso}, we derive the following description of all polytopal realizations of the fan~$\gvectorFan[n]$, recovering all associahedra of~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan}.
Note that the arguments of~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} were quite different from the present~approach.

@@

\para{Non-kissing fan and gentle associahedron}
%
\enlargethispage{.1cm}
The $\b{g}$-vectors support a complete simplicial fan realization of the non-kissing complex~$\NKC$.
Examples are illustrated in \cref{fig:nonkissingFans}.

@@

It is proved in~\cite[Thm.~4.27]{PaluPilaudPlamondon-nonkissing} that the non-kissing fan comes from a polytope.
For a walk~$\omega$, denote by~$\KN(\omega)$ the sum over all other walks~$\omega'$ of the number of kisses between~$\omega$~and~$\omega'$.

@@

\enlargethispage{-.3cm}
We now discuss the type cones of the non-kissing fans defined in \cref{thm:nonkissingFan}. We obtain from \cref{prop:exchangeablePairsNKC} the following redundant description.

@@

%\newpage
\part{Grothendieck group and relations between $\b{g}$-vectors}
\label{part:algebra}

>>>>

Our assumptions allow for a well-defined notion of index:~$\ind_T(X)=[T_0^X]-[T_1^X] \in \kzero{\add T}$.

@@

Fix an essential complete simplicial fan~$\Fan$ in~$\R^n$, with an (arbitrary) indexing of its rays by $[N] \eqdef \{1,\dots,N\}$. Let~$\b{G}$ be the $N \times n$-matrix whose rows are representative vectors of the rays of~$\Fan$. Let~$\b{K}$ be a $(N-n) \times N$-matrix that spans the left kernel of~$\b{G}$ (\ie $\b{K}\b{G} = 0$ and $\rank(\b{K})=N-n$). For any height vector~$\b{h} \in \R^N$, we define the polytope
\[
P_\b{h} \eqdef \set{\b{x} \in \R^n}{\b{G}\b{x} \le \b{h}},
\]
where $\b{G}\b{x} \le \b{h}$ is the standard shorthand for the corresponding system of inequalities.

@@

\begin{example}
\label{exm:typeCone}
\begin{figure}[b]
	\capstart
	\centerline{{\includegraphics[scale=1]{typeConeFan}} \quad {\includegraphics[scale=1]{typeConePolytope}} \quad \raisebox{.5cm}{\includegraphics[scale=.45]{typeCone}}}
	\caption{A $2$-dimensional fan~$\Fan$ with five rays $1, \dots, 5$ and five maximal cones~$a, \dots, e$ (left), its polytopal realization corresponding to the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, 1, 1, \nicefrac{5}{4})$ (middle), and the intersection of $\b{K}\typeCone(\Fan)$ with a hyperplane to get a $2$-dimensional slice (right).}
	\label{fig:typeCone}
\end{figure}
%
Consider the $2$-dimensional fan~$\Fan$ depicted in \cref{fig:typeCone} (left) (we will encounter this fan again in \cref{exm:typeConeCoarsenings,exm:typeConeMinkowskiSum,exm:typeConeAsso}).
It has five rays labeled~$1, \dots, 5$ and five maximal cones labeled~$a,\dots, e$.
For the matrices $\b{G}$ and $\b{K}$, we consider
\[
%\transpose{\b{G}} = \begin{bmatrix} 1 & 0 & -1 & -1 & 0 \\ 0 & 1 & 1 & 0 & -1 \end{bmatrix}
\b{G} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \\ -1 & 1 \\ -1 & 0 \\ 0 & -1 \end{bmatrix}
\qquad\text{and}\qquad
\b{K} = \begin{bmatrix} 1 & 0 & 0 & 1 & 0 \\ 0 & 1 & 0 & 0 & 1 \\ 1 & 0 & 1 & 0 & 1 \end{bmatrix}.
\]
The type cone~$\typeCone(\Fan)$ lies in~$\R^5$, but has a $2$-dimensional lineality space.
The five pairs of adjacent maximal cones of~$\Fan$ give rise to following five defining inequalities for~$\typeCone(\Fan)$:
\begin{gather*}
H^>_{ab}: \; \b{h}_1 + \b{h}_3 - \b{h}_2 > 0
\qquad
H^>_{bc}: \; \b{h}_2 + \b{h}_4 - \b{h}_3 > 0
\qquad
H^>_{cd}: \; \b{h}_3 + \b{h}_5 - \b{h}_4 > 0
\\
H^>_{de}: \; \b{h}_1 + \b{h}_4 > 0
\qquad
H^>_{ae}: \; \b{h}_2 + \b{h}_5 > 0,
\end{gather*}
where $H^>_{ij}$ denotes the halfspace defined by the inequality corresponding to the two adjacent maximal cones~$i$ and~$j$.
Note that the inequalities~$H^>_{ae}$ and~$H^>_{de}$ are redundant.
For example, the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$ belongs to the type cone~$\typeCone(\Fan)$, and the corresponding polytope is represented in \cref{fig:typeCone} (middle).
To represent this type cone, it suffices to represent~$\b{K}\typeCone(\Fan)$ which is an essential simplicial cone in~$\R^3$, given as the intersection of the following five open halfspaces:
\begin{gather*}
\b{K}H^>_{ab}: \; {x} - {y} +2 {z} > 0
\qquad
\b{K}H^>_{bc}: \; {x} + {y} - {z} > 0
\qquad
\b{K}H^>_{cd}: \; -{x} + {y} +2 {z} > 0
\\
\b{K}H^>_{de}: \; 2{x} + {z} > 0
\qquad
\b{K}H^>_{ae}: \; 2{y} + {z} > 0.
\end{gather*}
One can further reduce the dimension by intersecting with a transversal hyperplane to get a $2$-dimensional representation.
This is the red triangle depicted in \cref{fig:typeCone} (right).
\end{example}

@@

We now discuss the connection between convex combinations in the type cone and Minkowski sums of polytopes.
Recall that the \defn{Minkowski sum} of two polytopes~$P$ and~$Q$ is the polytope $P + Q \eqdef \set{p + q}{p \in P\text{ and } q \in Q}$.

@@

\begin{example}
\label{exm:typeConeMinkowskiSum}
Consider the $2$-dimensional fan~$\Fan$ of \cref{fig:typeCone}\,(left) whose type cone has been described in \cref{exm:typeCone} and represented in \cref{fig:typeCone}\,(right).
As illustrated in \cref{fig:typeConeCoarsenings}, the rays of the closed type cone~$\ctypeCone(\Fan)$ are directed by the height vectors~$(1,2,1,2,1)$ (corresponding to a triangle), $(0,1,1,0,1)$ (corresponding to a vertical segment) and~$(1,0,1,1,0)$ (corresponding to a horizontal segment).
Therefore, any polytope in~$\ctypeCone(\Fan)$ is (up to translation) a positive Minkowski combination of a triangle and two edges.
\cref{fig:typeConeMinkowskiSum} illustrates this property for the polytope of \cref{fig:typeCone}\,(middle) corresponding to the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$.
Note that~$\Fan$ is the normal fan of J.-L.~Loday's $2$-dimensional associahedron, whose realization as a Minkowski sum of faces of a simplex was given by A.~Postnikov in~\cite{Postnikov}.

\begin{figure}
	\capstart
	\centerline{
	\begin{tabular}{c@{\;$=$\;}c@{\,$+$\,}c@{\,$+$\,}c@{\,}l@{\,}l}
	$(\frac{1}{2}, \frac{3}{4}, \frac{5}{4}, 1, \frac{5}{4})$ & $\frac{1}{6}(1,2,1,2,1)$ & $\frac{3}{4}(0,1,1,0,1)$ & $\frac{1}{2}(1,0,1,1,0)$ & $-\,\frac{1}{6}(1,0,-1,-1,0)$ & $-\,\frac{1}{3}(0,1,1,0,-1)$ \\
	\raisebox{-0.5\height}{\includegraphics[scale=.5]{typeConePolytope}} & \raisebox{-0.5\height}{\includegraphics[scale=.5]{typeConeMinkowskiSum1}} & \raisebox{-0.5\height}{\includegraphics[scale=.5]{typeConeMinkowskiSum2}} & \raisebox{-0.5\height}{\includegraphics[scale=.5]{typeConeMinkowskiSum3}} & $+$ translation.
	\end{tabular}
	}
	\caption{The polytope of \cref{fig:typeCone}\,(middle) is a Minkowski sum of a triangle and two edges.}
	\label{fig:typeConeMinkowskiSum}
\end{figure}
\end{example}

@@

\pagebreak
We now focus on the normal fan of~$\Asso[n]$.
Since~$\Asso[n]$ lives in a hyperplane of~$\R^{n+1}$, its normal fan has a one-dimensional lineality.
Let~$\HH \eqdef \bigset{\b{x} \in \R^{n+1}}{\sum_{\ell \in [n+1]} \b{x}_\ell = 0}$ and ${\pi : \R^{n+1} \to \HH}$ denote the orthogonal projection. We denote by~$\gvectorFan[n]$ the intersection of the normal fan of~$\Asso[n]$ with~$\HH$, which is an essential complete simplicial fan.

@@

\cref{thm:normalFanLoday} is illustrated in \cref{fig:lodayFans} and \cref{exm:typeConeAsso} in dimension~$2$ and~$3$.
Note that in \cref{fig:lodayFans} and \cref{exm:typeConeAsso} we express the $\b{g}$-vectors in the basis given by $\gvector{0,i+1}$ for $i\in [n]$ (to be coherent with the upcoming \cref{def:gvectorCA}).

\begin{figure}
	\capstart
	\centerline{\raisebox{.5cm}{\includegraphics[scale=.45]{lodayFan2}}\hspace{1.5cm}\includegraphics[scale=.45]{lodayFan3}}
	\caption{The normal fans~$\gvectorFan[2]$ and~$\gvectorFan[3]$ of J.-L.~Loday's associahedra. As the rightmost fan is $3$-dimensional, we intersect it with the sphere and stereographically project it from the direction~$(-1,-1,-1)$.}
	\label{fig:lodayFans}
\end{figure}

@@

From the linear dependencies of \cref{prop:exchangeablePairsAsso}, we obtain a redundant description of the type cone of the fan~$\gvectorFan[n]$.
To simplify the presentation and write the wall-crossing inequalities in a uniform way, we embed the type cone in a larger space, adding dummy variables for the boundary edges of the polygon.

@@

\pagebreak

@@

\begin{figure}[h]
	\capstart
    \begin{adjustbox}{center}
        \begin{tikzpicture}
        	\matrix (m) [matrix of math nodes, row sep=.35cm, column sep=.2cm, nodes={anchor=center, align=center, inner sep=0pt}]{
        		& \;\includegraphics[scale=.7]{diagonalA2} & \node (b) {\red \circled{B}}; & \;\includegraphics[scale=.7]{diagonalA4} & \\
        		\;\includegraphics[scale=.7]{diagonalA1} & \node (a) {\red \circled{A}}; & \;\includegraphics[scale=.7]{diagonalA3} & \node (c) {\red \circled{C}}; & \;\includegraphics[scale=.7]{diagonalA5} \\};
        	\draw[->] (m-2-1) -- (m-1-2);
        	\draw[->] (m-1-2) -- (m-2-3);
        	\draw[->] (m-2-3) -- (m-1-4);
        	\draw[->] (m-1-4) -- (m-2-5);
        	\draw[red] (a) -- (m-2-1);
        	\draw[red] (a) -- (m-2-3);
        	\draw[red] (a) -- (m-1-2);
        	\draw[red] (b) -- (m-1-2);
        	\draw[red] (b) -- (m-1-4);
        	\draw[red] (b) -- (m-2-3);
        	\draw[red] (c) -- (m-2-3);
        	\draw[red] (c) -- (m-2-5);
        	\draw[red] (c) -- (m-1-4);
        \end{tikzpicture}
        \quad
        \begin{tikzpicture}
        	\matrix (m) [matrix of math nodes, row sep=.35cm, column sep=.2cm, nodes={anchor=center, align=center, inner sep=0pt}]{
        		&& \;\includegraphics[scale=.7]{diagonalB3} & \node (c) {\red \circled{C}}; & \;\includegraphics[scale=.7]{diagonalB7} && \\
        		& \;\includegraphics[scale=.7]{diagonalB2} & \node (b) {\red \circled{B}}; & \;\includegraphics[scale=.7]{diagonalB5} & \node (e) {\red \circled{E}}; & \;\includegraphics[scale=.7]{diagonalB8} & \\
        		\;\includegraphics[scale=.7]{diagonalB1} & \node (a) {\red \circled{A}}; & \;\includegraphics[scale=.7]{diagonalB4} & \node (d) {\red \circled{D}}; & \;\includegraphics[scale=.7]{diagonalB6} & \node (f) {\red \circled{F}}; & \;\includegraphics[scale=.7]{diagonalB9} \\};
        	\draw[->] (m-3-1) -- (m-2-2);
        	\draw[->] (m-2-2) -- (m-1-3);
        	\draw[->] (m-2-2) -- (m-3-3);
        	\draw[->] (m-1-3) -- (m-2-4);
        	\draw[->] (m-3-3) -- (m-2-4);
        	\draw[->] (m-2-4) -- (m-1-5);
        	\draw[->] (m-2-4) -- (m-3-5);
        	\draw[->] (m-1-5) -- (m-2-6);
        	\draw[->] (m-3-5) -- (m-2-6);
        	\draw[->] (m-2-6) -- (m-3-7);
        	\draw[red] (a) -- (m-3-1);
        	\draw[red] (a) -- (m-3-3);
        	\draw[red] (a) -- (m-2-2);
        	\draw[red] (b) -- (m-2-2);
        	\draw[red] (b) -- (m-2-4);
        	\draw[red] (b) -- (m-1-3);
        	\draw[red] (b) -- (m-3-3);
        	\draw[red] (c) -- (m-1-3);
        	\draw[red] (c) -- (m-1-5);
        	\draw[red] (c) -- (m-2-4);
        	\draw[red] (d) -- (m-3-3);
        	\draw[red] (d) -- (m-3-5);
        	\draw[red] (d) -- (m-2-4);
        	\draw[red] (e) -- (m-2-4);
        	\draw[red] (e) -- (m-2-6);
        	\draw[red] (e) -- (m-1-5);
        	\draw[red] (e) -- (m-3-5);
        	\draw[red] (f) -- (m-3-5);
        	\draw[red] (f) -- (m-3-7);
        	\draw[red] (f) -- (m-2-6);
        \end{tikzpicture}
    \end{adjustbox}
	\caption{The facet-defining inequalities of the type cone~$\typeCone \big( \gvectorFan[n] \big)$ correspond to the flips described in \cref{prop:extremalExchangeablePairsAsso}. See \cref{subsec:typeConeCA,sec:clusterCategories} for a representation theoretic \mbox{interpretation}.}
	\label{fig:labelFacetDefiningInequalititiesAsso}
\end{figure}

@@

\[
\begin{array}{r|cccccccccc}
\text{diagonals} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB1}}  & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB2}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB3}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB4}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB5}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB6}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB7}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB8}} & \raisebox{-.25cm}{\;\includegraphics[scale=.7]{diagonalB9}} \\[.3cm]
\text{$\b{g}$-vectors} & \compactVectorT{1}{0}{0} & \compactVectorT{0}{1}{0} & \compactVectorT{0}{0}{1} & \compactVectorT{-1}{1}{0} & \compactVectorT{0}{-1}{1} & \compactVectorT{-1}{0}{1} & \compactVectorT{-1}{0}{0} & \compactVectorT{0}{-1}{0} & \compactVectorT{0}{0}{-1} \\[.6cm]
\text{facet}		& 1 & -1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & \red \circled{A} \\
\text{defining} 	& 0 & 1 & -1 & -1 & 1 & 0 & 0 & 0 & 0 & \red \circled{B} \\
\text{inequalities}	& 0 & 0 & 1 & 0 & -1 & 0 & 1 & 0 & 0 & \red \circled{C} \\
					& 0 & 0 & 0 & 1 & -1 & 1 & 0 & 0 & 0 & \red \circled{D} \\
					& 0 & 0 & 0 & 0 & 1 & -1 & -1 & 1 & 0 & \red \circled{E} \\
					& 0 & 0 & 0 & 0 & 0 & 1 & 0 & -1 & 1 & \red \circled{F} \\[.2cm]
\end{array}
\]

@@

\pagebreak

@@

Combining \cref{coro:simplicialTypeCone,coro:simplicialTypeConeAsso} and \cref{prop:extremalExchangeablePairsAsso}, we derive the following description of all polytopal realizations of the fan~$\gvectorFan[n]$, recovering all associahedra of~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan}.
Note that the arguments of~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} were quite different from the present~approach.

@@

\para{Non-kissing fan and gentle associahedron}
%
The $\b{g}$-vectors support a complete simplicial fan realization of the non-kissing complex~$\NKC$.
Examples are illustrated in \cref{fig:nonkissingFans}.

@@

\enlargethispage{.1cm}
It is proved in~\cite[Thm.~4.27]{PaluPilaudPlamondon-nonkissing} that the non-kissing fan comes from a polytope.
For a walk~$\omega$, denote by~$\KN(\omega)$ the sum over all other walks~$\omega'$ of the number of kisses between~$\omega$~and~$\omega'$.

@@

We now discuss the type cones of the non-kissing fans defined in \cref{thm:nonkissingFan}. We obtain from \cref{prop:exchangeablePairsNKC} the following redundant description.

@@

\newpage
\part{Grothendieck group and relations between $\b{g}$-vectors}
\label{part:algebra}
